Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-13 paper.tex
wordings in section 2

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
% Lars: "orders of magnitude faster" is an implementation detail
on all NumPy and SciPy vector operations while allowing us at the same time to
keep the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices, hence helping them to familiarize with the
project.  From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

The public API is oriented towards batch processing rather than processing
single samples per API call. While classification and regression algorithms can
indeed make predictions for single samples, \sklearn objects are not optimized
for this use case. (The few online learning algorithms implemented are intended
to take minibatches.) Batch processing makes optimal use of NumPy and SciPy by
preventing the overhead inherent to Python function calls or due to per-element
dynamic type checking. Although this might seem to be an artifact of the Python
language, and therefore an implementation detail that leaks into the API, we
argue that APIs should be designed so as not to tie a library to a suboptimal
implementation strategy. As such, batch processing enable fast implementations
in lower-level languages (where memory hierarchy effects and the possibility of
internal parallelization come into play).

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as inputs a dataset \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in the case of classification or the mean squared
error in the case of regression. Not however that, from an API point of view,
the only guideline is that the \texttt{score} method returns a numerical value
such that the higher the better. As such, some unsupervised learning estimators
also exposes a \texttt{score} function, computing for instance the likelihood of
the given data.

>>>>

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
% Lars: "orders of magnitude faster" is an implementation detail
on all NumPy and SciPy vector operations while allowing us at the same time to
keep the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
% Gilles: the phrase was too long
% hence helping them to familiarize with the project.
From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

For efficiency reasons, the public interface is oriented towards processing
batches of samples rather than single samples per API call. While classification
and regression algorithms can indeed make predictions for single samples,
\sklearn objects are not optimized for this use case. (The few online learning
algorithms implemented are intended to take mini-batches.) Batch processing makes
optimal use of NumPy and SciPy by preventing the overhead inherent to Python
function calls or due to per-element dynamic type checking. Although this might
seem to be an artifact of the Python language, and therefore an implementation
detail that leaks into the API, we argue that APIs should be designed so as not
to tie a library to a suboptimal implementation strategy. As such, batch
processing enable fast implementations in lower-level languages (where memory
hierarchy effects and the possibility of internal parallelization come into
play).

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in the case of classification, or the mean squared
error in the case of regression. Not however that, from an API point of view,
the only guideline is that the \texttt{score} method returns a numerical value
such that the higher the better. As such, some unsupervised learning estimators
also exposes a \texttt{score} function, computing for instance the likelihood of
the given data.
